%=============================================================================
% Chapter: Code Organization
%=============================================================================

\chapter{Code Organization}
\label{Code Organization}

In this chapter, we describe various organizational issues
regarding the \parflow{} code.

%=============================================================================
%=============================================================================

\section{The Developer's Environment}
\label{The Developer's Environment}

This section assumes that the developer is working on the workstation
cluster at LLNL.  No procedures have yet been devised for developers
working elsewhere (there currently are no such developers).

The \parflow{} repository (central location for all of the \parflow{}
source) is located in \file{/home/casc/repository/parflow}.  Online
documentation, WWW information, and data for many \parflow{} runs, are
all located in a \kbd{parflow} account (directory \file{~parflow}) on
the local workstation cluster.  Developers work on their own local
versions of the code, which gets ``checked'' in and out of the
repository as changes are made.  The source code control system that
is being used is the CVS system.  The following steps describe how to
check out a ``clean'' version of the code.

\begin{enumerate}

%------------------------------
\item
The following environment variables should be set up in your \file{.profile}
file (or similar shell resource file):

\begin{display}\begin{verbatim}
CVSROOT=/home/casc/repository
PARFLOW_SRC=~/parflow/src
PARFLOW_DIR=~/parflow/exe
PARFLOW_HELP=~parflow/docs
export CVSROOT PARFLOW_SRC PARFLOW_DIR PARFLOW_HELP

\end{verbatim}\end{display}

You also need to add \code{\$PARFLOW_DIR/bin} to your 
\code{\$PATH} environment variable.

The variable \code{CVSROOT} specifies the location of the parflow
repository and other \code{CVS} repositories.
The variable \code{PARFLOW_DIR} specifies the location of
the installed version of \parflow{}.
To allow more than one binary version of the code to be installed in
your directory at a time, set
\begin{display}\begin{verbatim}
PARFLOW_DIR=~/parflow/exe.`uname`
\end{verbatim}\end{display}
This will, for example,
install Sun binaries in \file{~/parflow/exe.SunOS}
and SGI binaries in \file{~/parflow/exe.IRIX}.
\code{PARFLOW_SRC} is the location of the source code for \parflow{}
and affiliated tools.
\code{PARFLOW_HELP} is the location of the online help files.
We will use the \file{~/parflow} directory as the root directory for
building \parflow{} in this manual; you may use a different directory
if you wish.


%------------------------------
\item
Get the source code out of the repository and run the post checkout
script.

\begin{display}\begin{verbatim}
mkdir ~/parflow
cvs checkout -P parflow
(cd ~/parflow/; ./post_checkout)
\end{verbatim}\end{display}

This creates a local \parflow{} directory tree from the sources in 
the repository.

%------------------------------
\item
To build and run the code, refer to the {\em ParFlow User's Manual}.

\end{enumerate}

%=============================================================================
%=============================================================================

\section{Building Code}
\label{Building Code}

All code and documentation are ``built'' with a \file{build} script.
When writing new \file{build} scripts, the following \file{build}
command conventions should be followed:

\begin{itemize}
\item
\kbd{build all} --- builds things
\item
\kbd{build install} --- builds and installs exe's and support files.
Installation usually involves copying of files into the
\code{\$PARFLOW_DIR} and \code{\$PARFLOW_HELP} directories.
However, in some cases, files are installed elsewhere.
\item
\kbd{build clean} --- clean object files (and some other temp stuff)
\item
\kbd{build veryclean} --- cleans object files and exe's within
the \code{\$PARFLOW_SRC} directory (i.e., does not remove ``installed''
files).
\end{itemize}

This build technology contains some ``smarts''.
For example, it detects when you change platforms (say you log into the
SGI) and will delete object files and remake everything.

\noindent
{\bf Notes:}

\begin{itemize}
\item
Because code is ``copied'' when installed (as opposed to using
soft links, for example), when changes are made to any source,
help files, input files, etc., you {\em must} do a \kbd{build install}
to update the changes in the ``install'' directory.

\item
You can safely ignore this item if you run parflow out of your
home directory.
Also, I believe that this only impacts the Chameleon port so you
can probably safely ignore all this.

\noindent
Due to some NFS temp mount problems there are some limitations to how
we can run things.
The problem is attempting to get the ``virtual'' current directory
rather than the ``real'' directory name.
For example, if you do a \kbd{pwd}, you will sometimes see
\file{/home/mercury0/ssmith/blah/foo} or
\file{/export/home/mercury0/ssmith/blah/foo}
(if you are on the machine where your home directory is),
and other times you will see 
\file{/tmp_mnt/home/mercury0/ssmith/blah/foo}
(if you are on a remote machine).
This is a problem when running some of the different
models since they need to see the {\em same} directory
(\file{/home/mercury0/ssmith/blah/foo} is the ``correct'' name).
To handle this, everything is based off of the \code{AMPS_ROOT_DIR}
environment variable.
By default this is set to your home directory.
If the directory that you are running from is in your \code{\$HOME}
directory, then you are fine.
If you want to run from somewhere else (say \file{/scratch/parflow/exe})
you need to either put a softlink in your home directory or explicitly set
\code{AMPS_ROOT_DIR} to \file{/scratch}.
You do {\em not} have to worry about this if you run \parflow{} out of
your home directory.

\end{itemize}

%=============================================================================
%=============================================================================

\section{Checking out code}
\label{Checking out code}

Use the \code{cvs edit} command to check code out from the \parflow{} source
code repository.

%=============================================================================
%=============================================================================

\section{Checking in code}
\label{Checking in code}

Use the \code{cvs commit} utility to check in your changes to the
\parflow{} repository.  If you add new code you will have to use
the \code{cvs add} to add it.

%=============================================================================
%=============================================================================

\section{Installing a new version of ParFlow}
\label{Installing a new version of ParFlow}

This section discusses how to install a new version of \parflow 
for general users; this is referred to by the \parflow team as the
``installed'' version of \parflow.

This will describe the steps need to install the version on the CASC
machines (which currently consist of the Sparc Cluster and the SGI
Onyx).

The first thing to do is install on the Sparc Cluster:

\begin{enumerate}
\item
Log in to the parflow account on the Sparc Sun Cluster
\item
Execute the script to build and install parflow.  This will take 
a rather long time.
\begin{display}\begin{verbatim}
(cd /home/casc/parflow/;./install.SunOS)
\end{verbatim}\end{display}
\item
Execute the test program to make sure things are running OK
\begin{display}\begin{verbatim}
(cd /home/casc/parflow/exe.`uname`/test/; tclsh default\_single.pftcl)
\end{verbatim}\end{display}
\item
ParFlow should run on the default\_single problem
\end{enumerate}

If that worked, install on the SGI:

\begin{enumerate}
\item
Log into nyx (as the parflow user)
\begin{display}\begin{verbatim}
rlogin nyx
\end{verbatim}\end{display}
\item
Build and install parflow
\begin{display}\begin{verbatim}
(cd /home/casc/parflow/; ./install.IRIX64)
\end{verbatim}\end{display}
\item
Execute the test program to make sure things are running OK
\begin{display}\begin{verbatim}
(cd /home/casc/parflow/exe.`uname`/test/; tclsh default\_single.pftcl)
\end{verbatim}\end{display}
\item
ParFlow should run
\end{enumerate}

Things should be good to go.  Email ``pfusers'' to tell everyone
that you have installed a new version of parflow.

===========================================================================



