%=============================================================================
% Chapter: Gridding
%=============================================================================

\chapter{Gridding}
\label{Gridding}

The Applied Mathematics Group has extensive experience with adaptive
mesh refinement techniques, or {\em AMR}.
In an effort to utilize this knowledge, and keep coherent the various
efforts within the {\em CCSE}, the gridding within \parflow{} is
a parallel adaptation of {\em AMR}.
The differences between the two approaches are mostly a consequence
of parallel computing.

See also, \ref{Grid Reference}.

%=============================================================================
% Section: Terminology
%=============================================================================

\section{Terminology}
\label{Terminology}

The \dfn{background} is a uniform grid with position in
real-space given by the quantities
$x_0$, $y_0$, $z_0$, $N_x$, $N_y$, $N_z$, $\Delta x$, $\Delta y$, $\Delta z$
(see Figure~\ref{fig-grid}).

An \dfn{index-space} defines a specific correspondence between
integer index triples and real-space coordinates.
This correspondence is defined in terms of the background
(for index-space $l$)
by the resolutions, $(r_{l,x}, r_{l,y}, r_{l,z})$,
and an alignment convention (which would differ for cell-centered
gridding; we are doing vertex-centered gridding at this time).

Specifically, the real-space coordinate, $(x_{i,l}, y_{j,l}, z_{k,l})$,
associated with the integer index triple, $(i, j, k)$,
on index-space $l$ is given by:
\begin{eqnarray}
x_{i,l} & = & x_0 + i \left( {{\Delta x} \over {2^{r_{l,x}}}} \right) \\
y_{j,l} & = & y_0 + j \left( {{\Delta y} \over {2^{r_{l,y}}}} \right) \\
z_{k,l} & = & z_0 + k \left( {{\Delta z} \over {2^{r_{l,z}}}} \right)
\end{eqnarray}
Note: the resolutions may be negative, indicating coarser spacing.
Each index-space is associated with a unique \dfn{refinement level}.
These levels are labeled by the sum $r_{l,x} + r_{l,y} + r_{l,z}$.
Note that since levels are unique, this means that for any two
levels, $l$ and $m$,
\begin{eqnarray}
r_{l,x} & >= & r_{m,x}, \\
r_{l,y} & >= & r_{m,y}, \\
r_{l,z} & >= & r_{m,z},
\end{eqnarray}
where strict inequality must hold for at least one of the above
equations.

A \dfn{subregion} is a collection of indices on index-space
$l$, described by the quantities
$i_{l,x}$, $i_{l,y}$, $i_{l,z}$,
$n_{l,x}$, $n_{l,y}$, $n_{l,z}$,
$s_{l,x}$, $s_{l,y}$, $s_{l,z}$.
Specifically, a subregion is defined by
(see Figure~\ref{fig-grid})
\begin{eqnarray}
& & \{ i_{l,x} + i s_{l,x} : i = 0, 1, \ldots, (n_{l,x}-1) \} \otimes \\
& & \{ i_{l,y} + i s_{l,y} : i = 0, 1, \ldots, (n_{l,y}-1) \} \otimes \\
& & \{ i_{l,z} + i s_{l,z} : i = 0, 1, \ldots, (n_{l,z}-1) \} .
\end{eqnarray}
The striding factors allow one to define things
like ``red points'' or ``black points'' for red/black iterative methods,
or ``coarse points'' and ``fine points'' for multigrid methods.

A \dfn{subregion-array} is just an array of subregions.

A \dfn{region} is an array of subregion-arrays.

A \dfn{subgrid} is a subregion with striding factors equal 1.

A \dfn{subgrid-array} is an array of subgrids.

A \dfn{grid} is also an array of subgrids.

%=============================================================================
% Section: Gridding Requirements
%=============================================================================

\section{Gridding Requirements}
\label{Gridding Requirements}

\begin{itemize}

\item
The union of subgrids at level 0 must cover the entire problem domain.
The union of subgrids at a refined level $l > 0$
may only cover a portion of the problem domain.
 
\item
Subgrids may overlap.
 
\item
The computed solution at a point $P$
in the problem domain
is defined to be the value at the point on the finest subgrid containing it.
(note: we use ``vertex centered'' gridding, not ``cell centered'')

\item
The region of the domain covered by level $m > l$ subgrids
must be contained in the region covered by level $l$ subgrids.
This containment is not required to be proper (since we have vertex
centered grids).
 
\end{itemize}

