%=============================================================================
% Chapter: ParFlow Input
%=============================================================================

\chapter{ParFlow Input}
\label{ParFlow Input}

Input into \parflow{} is done using a TCL script and a simple 
database system.  The database system is based on key/value pairs.
The user sets values in the input file using
the \code{pfset} command.  When the \code{pfrun} routine is executed
it creates a \file{runname.pfidb} file which contains the key/value
database.  Inside the \parflow{} code the database is read in using the
\code{IDB_NewDB} command.  The database is then queried using a set
of macros \code{GetType} and \code{GetTypeDefault} where \code{Type}
is one of String, Int, or Double.  All of these routines take a key
value as input.  The default versions also take a default value (of
the appropriate type).

\begin{display}\begin{verbatim}
   char *geom_name;

   geom_name = GetString("Domain.GeomName");

   geom_name = GetStringDefault("Domain.GeomName", "DefaultGeom");

\end{verbatim}\end{display}

The input is divided into two logical sections.  The first is general
global problem setup.  This contains items like the geounits.  The
other section is used to set up the \parflow{} modules.  Each of the
modules has a name which is based on the call tree.  It starts with
``Solver'' and each module appends names to this.  In this way each module
instance can have it's own input.  A module retrieves module specific
values based on the name that is passed into the NewPublicXtra routine.

\begin{display}\begin{verbatim}

/* name is the name that this module was given by it's parent */
PFModule   *FooNewPublicXtra(char *name)
{
   char key[IDB_MAX_KEY_LEN];   /* key name use to query DB */
        

   sprintf(key, "%s.MaxLevels", name); /* construct the key name from
                                        the modules name and the value
                                        we need */
   public_xtra -> max_levels = GetIntDefault(key, 100);
}
\end{verbatim}\end{display}


When a module instantiates a module which it is going to invoke it
passes in it's module name with an addition name appended to.  The
parent modules name their children.  For example an iterative solver
would pass into a preconditioner MyName.Precond (where ``MyName''
would be what this module was given as it's name).  This constructs a
tree like naming convention to ensure that each module instance has a
unique name to use as a key for getting input that is specific to that
module instance.

In some cases you want to provide a choice on which module should be used.
For example you might wish to let the user choose among several different
preconditioners.  This is done by querying the database for the name
of the preconditioner to invoke and using a standard switch statement to
invoke the appropriate \code{NewModule} routine.

\begin{display}\begin{verbatim}
   char key[IDB_MAX_KEY_LEN];

   char          *switch_name;
   int            switch_value;

   /* Name arrays are helper structures to parse space separated strings */
   /* A name array can convert from an index to name or from a name      */
   /* to an index.                                                       */
   NameArray      switch_na;

	
   /* Build the name array with two names:                               */
   /*   0 = RedBlackGSPoint                                              */
   /*   1 = WJacobi                                                      */
   switch_na = NA_NewNameArray("RedBlackGSPoint WJacobi");

   sprintf(key, "%s.Smoother", name);
   switch_name = GetStringDefault(key,"RedBlackGSPoint");

   /* Convert the name the user entered into an int to use in the switch */
   /* -1 is returned if the name is not found in the name array          */
   switch_value  = NA_NameToIndex(switch_na, switch_name);
   switch (switch_value)
   {
      case 0:
      {
	 /* key has the name we want to give to this child */
         public_xtra -> smooth = PFModuleNewModule(RedBlackGSPoint, (key));
         break;
      }
      case 1:
      {
         public_xtra -> smooth = PFModuleNewModule(WJacobi, (key));
         break;
      }
      default:
      {
         InputError("Error: Invalid value <%s> for key <%s>\n", switch_name,
                     key);
      }
   }
   NA_FreeNameArray(switch_na);

}
\end{verbatim}\end{display}
