%=============================================================================
% Chapter: Visualization
%=============================================================================

\chapter{Visualization}
\label{Visualization}

This chapter discusses visualization of \parflow{} data.
In \S~\ref{AVS} we discuss AVS,
in \S~\ref{AVS/Express} we discuss AVS/Express, and
in \S~\ref{Explorer}, we discuss Explorer.

%=============================================================================
% Section: AVS
%=============================================================================

\section{AVS}
\label{AVS}

This section discusses the AVS visualization system.
In \S~\ref{Writing AVS Modules} we describe how to write AVS modules, and
in \S~\ref{Writing AVS Coroutines} we describe how to write AVS coroutines.

%=============================================================================
% SubSection: Writing AVS Modules
%=============================================================================

\subsection{Writing AVS Modules}
\label{Writing AVS Modules}

Before continuing, see the note in \S~\ref{Writing AVS Module Help}
on naming conventions.
Now suppose we want to create the module ``foo~bar''.
We then do the following:
\begin{enumerate}

\item
In file \file{foo_bar.c}, write the functions \code{FooBar} and
\code{FooBar_compute}.
Use the examples in the \file{avs/modules} directory as a guide.

\item
In \file{pf_modules.c}, add the lines:
\begin{display}\begin{verbatim}
int FooBar();
AVSmodule_from_desc(FooBar);
\end{verbatim}\end{display}

\item
In \file{Makefile}, add \code{foo_bar.c} to the \code{SRC} string definition.

\item
Type
\begin{display}\begin{verbatim}
build install
\end{verbatim}\end{display}
This installs the module library in the directory
\file{\$PARFLOW_DIR/avs/mod_lib}.

\item
Test the module library by using the \file{pfavs} script.

\end{enumerate}

%=============================================================================
% SubSection: Writing AVS Coroutines
%=============================================================================

\subsection{Writing AVS Coroutines}
\label{Writing AVS Coroutines}

Before continuing, see the note in \S~\ref{Writing AVS Module Help}
on naming conventions.
Now suppose we want to create the coroutine ``foo~bar''.
We then do the following:
\begin{enumerate}

\item
In file \file{foo_bar.c}, write the function \code{FooBar}
and a \code{main} routine.
Use the examples in the \file{avs/modules} directory as a guide.

\item
In \file{Makefile}, add \code{foo_bar} to the \code{COROUTINES}
string definition, and define a rule for compiling \code{foo_bar}
(use \code{rotate_geom} as an example).

\item
In \file{ParFlow_source}, add the line
\begin{display}\begin{verbatim}
file foo_bar
\end{verbatim}\end{display}

\item
Type
\begin{display}\begin{verbatim}
build install
\end{verbatim}\end{display}
This installs the coroutine in the directory
\file{\$PARFLOW_DIR/avs/mod_lib}.

\item
Test the coroutine by using the \file{pfavs} script.

\end{enumerate}


%=============================================================================
% Section: AVS/Express
%=============================================================================

\section{AVS/Express}
\label{AVS/Express}

This section discusses the AVS/Express visualization system.
In \S~\ref{Existing AVS/Express Modules} we describe the aspects of the
AVS/Express applications written for \parflow{} that are relevant to
developers.
In \S~\ref{Writing AVS/Express Modules} we describe how to write additional
modules and applications.

%=============================================================================
% SubSection: Existing AVS/Express Modules
%=============================================================================

\subsection{Existing AVS/Express Modules}
\label{Existing AVS/Express Modules}

There are a number of details relevant to the existing applications that
should be mentioned for the benefit of anyone who might modify the code.  The
first of these is about the structure of the applications.  Each is primarily
split up into three separate macros; one for reading in the data, one for the
user interface, and one for displaying the data.  This seems to be an
effective and logical way to organize the complexity of the networks.  It
should be noted though that the macros for reading the data and displaying it
provide many of the auxiliary user interface panels under the module stack.

Of the modules in the \menu{Library Workspaces}, the most useful to
developers of new applications will probably be \code{PF_Grid_from_File},
\code{Load_File_Frame}, \code{FileDialog_button}, and the modules in the
\code{string functions} library.  The \code{PF_Grid_from_File} module, which
will be mentioned later, contains some of the other modules in the library,
including the \code{read_parflow} module.  The \code{Load_File_Frame} module
is the top frame in all of the application user interfaces.  You should be
able to reuse it simply by putting it in your network and connecting it to
the UImod\_panel or other UI container.  The \code{FileDialog_button} module
is used to easily get a filename string into an application before a ``real''
user interface is built.  It simply displays a window with a button that
pops up a file dialog window when pressed.  The filename is the output of the
module.  The modules in the \code{string functions} library provide some
useful string manipulation operations.

The \code{PF_Grid_from_File} module
is the primary module used to read a \file{.pfb} or \file{.pfsb} file
into an AVS/Express field.  It should work ``out of the box'' and not need
modification in other networks, other than to change the default grid scaling
or downsizing.  The \code{merge_hack} and \code{Hack} groups in this module
are workarounds for bugs in the downsize module.  Without them, it is
impossible to get a field of type \code{Field_Unif} out of either the
\code{downsize} or \code{scale_grid} modules.  (The \code{Field_Unif} type is
needed for input to many of the modules in the applications).

It should be noted what is happening in the \code{PF_Grid_from_File} module
for \file{.pfsb} fields, as this is not totally straightforward.  After being
read into a structured field, they are converted to a uniform field of the
original grid size in the \code{sparse_to_unif} module.  The \code{ReadPFSB}
module has added an \code{orig_dims} array variable to the structured field,
so that the \code{sparse_to_unif} module knows how big to make the uniform
field.  If the \code{orig_dims} array is not present in the structured field,
\code{sparse_to_unif} prints a warning and assumes a $10\times10\times10$
grid.  This new uniform field is an input to the \code{Field_Chooser} module
which decides whether to output the \file{.pfb} or \file{.pfsb} field.

%=============================================================================
% SubSection: Writing AVS/Express Modules
%=============================================================================

\subsection{Writing AVS/Express Modules}
\label{Writing AVS/Express Modules}

%=============================================================================
% SubSubSection: Writing the Modules
%=============================================================================

\subsubsection{Writing the Modules}
\label{Writing the Modules}

Writing modules and applications for AVS/Express seems to be a good deal
easier than it is for AVS5.  It seems to require a lot less C code, and
creating user interfaces can even be done ``online.''  The best way to learn
how to write AVS/Express applications and modules is to read the {\it Getting
Started} manual and do the associated tasks in it.  This will take you
step-by-step through writing and running a small application.

%=============================================================================
% SubSubSection: Compiling AVS/Express Modules
%=============================================================================

\subsubsection{Compiling AVS/Express Modules}
\label{Compiling AVS/Express Modules}

The compilation process in AVS/Express compiles all source code specified
in the user library, so when you compile you are in effect compiling the
entire user library.
To compile from within Express, click on the library version of a module with
source code, and select \menu{Compile} from the \menu{Project} pulldown
menu.  (N.B. delete any active applications that contain your library modules
before doing so, or you will likely get an error).
To compile from the command prompt, use
\begin{display}\begin{verbatim}vxp -proj project_dir -compile -exit [-none]\end{verbatim}\end{display}
Using \kbd{-none} keeps Express from popping up any windows (you still need a
valid \code{DISPLAY} variable set though!).

This seems to work fine on
Solaris machines, but there seems to be bugs in the compilation process for
SGI machines.

To compile on an SGI you need to follow these steps:
\begin{enumerate}

\item Remove all the \code{-lsunmath} occurrences in the \file{v/templ.v}
file in the project directory.  Tip: run
\begin{display}\begin{verbatim}perl -pi.bak -e s/-lsunmath// v/templ.v\end{verbatim}\end{display}
in the project directory.  (This should be done {\em before}
loading Express).

\item Do one of the following.
    
    \begin{enumerate}
    \item Use the \menu{Compile} option within Express as described above,
	  ignoring the error message.
    \item From the command line run
\begin{display}\begin{verbatim}vxp -path "proj_dir global_express_dir" -generate -exit [-none]\end{verbatim}\end{display}
    \end{enumerate}

\item In the project directory run \kbd{make -f user.mk}.  This should
compile successfully despite a number of warnings.

\end{enumerate}

%=============================================================================
% SubSubSection: Tips and Tricks
%=============================================================================

\subsubsection{Tips and Tricks}
\label{Tips and Tricks}

If you are testing a module and it locks up Express, kill the \code{user}
process first, and if that does not fix it, kill the \code{express} process.
Often killing the user process will fix the situation and still leave
Express running, thus avoiding having to rerun it from the prompt.

To scroll the modules in the Network Editor, use shift-middle mouse button,
clicking on the background.  Use control-middle mouse button to scale the
modules.

If you are recompiling, and have not made major modifications to the
project (i.e. adding/deleting link files or source files, changing C
function names, etc.), you can save time by running \kbd{make -f user.mk} in
the project directory instead of using the \menu{Compile} option under the
\menu{Project} menu.


%=============================================================================
% SubSubSection: AVS/Express Bugs
%=============================================================================

\subsubsection{AVS/Express Bugs}
\label{AVS/Express Bugs}

AVS/Express 3.0 seems to have a good number of both bugs and ``unhelpful
features'' that are not quite bugs but act like them.  Outright bugs can be
found in the file \file{bugs.txt}, and some of the ``features'' are described
here.

In the \code{UIdial} module is one annoying feature.  If the \code{min}
variable becomes greater than or equal to the \code{max} variable,
\code{UIdial} resets one or both variables (usually only \code{min}) to some
outlandish value (like -99).  In doing so, it also breaks any connection that
the initial \code{min} (and/or \code{max}) variable upstream had.  This bites
you when you have the \code{min} and \code{max} values of the data from a
field connected to the \code{UIdial}.  If the \code{min} and \code{max}
values get out of sync, \code{UIdial} winds up breaking connections in the
field which automatically calculate the \code{min} and \code{max}.  Once this
is done, \code{min} or \code{max} become set to -99 or something similar, and
reloading the \parflow{} file does not reset it to the correct value.  The net
effect is that your display, \code{UIdial}, and colorbar all get hosed.  The
workaround:  use \code{copy_on_change} modules to protect the \code{min} and
\code{max} values of the field.  If you do it right, this protects everything
but the \code{UIdial} from getting hosed, and gives you some chance of
correcting the problem by reloading the \parflow{} file.  A better workaround:
create a group wrapper around the \code{UIdial} module, and write a method
which ensures that the \code{min} and \code{max} values being fed to the
\code{UIdial} module are in sync.

Being unable to connect an output port to an input port that look like they
should be connected is another common problem.  This is usually caused by one
of two things; either
\begin{itemize}
\item the input port already has a connection that is not visible, or
\item the output port type is not the same type or a supertype of the input port type.
\end{itemize}
These are not really unhelpful features so much as quirks of the AVS/Express
design and interface.  The first is easiest to correct; just open the input
port's variable and delete the contents, then try again to connect the two.
The second is generally more difficult.  First, try connecting the inputs of
the upstream module before connecting it to the downstream module.  Sometimes
this works when the output of the upstream module becomes the type of its
input.  If that doesn't work, you can sometimes hack something together using
the built-in merge operator to get the necessary components into the upstream
module output field.  (See the \code{merge_hack} and \code{Hack} modules in
\code{PF_Grid_from_File} for an example of this).

If modifying a library module does not produce the desired effect in an
application, it may be because the application is defining the module itself
instead of referencing the library module.  This can be easily checked by
looking for the module in the \file{.v} file for the application.  It will say
something like \code{group modulename} or \code{module modulename} if the
module is defined in the file.

See also the Common Problems section, 6.2, in {\it Visualizing Your Data with
AVS/Express} for other problems.  It's not great, but does give some useful
solutions.


%=============================================================================
% Section: Explorer
%=============================================================================

\section{Explorer}
\label{Explorer}

%=============================================================================
% SubSection: Writing Explorer Modules
%=============================================================================

\subsection{Writing Explorer Modules}
\label{Writing Explorer Modules}

