\documentstyle[11pt,makeidx,html,parflow]{book}

%=============================================================================
% Preamble:
%=============================================================================

% set margins
\setlength{\oddsidemargin}{0in} \setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.50in} \setlength{\topmargin}{0in}
\setlength{\textheight}{8.0in}

% define various commands and macros
\input{defs}
\input{inputps}

% define the version number
\input{version}

% write out the index entries in the `.idx' file
\makeindex

%=============================================================================
% Body:
%=============================================================================

\begin{document}

%=============== Title Page

\begin{TitlePage}

\Title{\parflow{} User's Manual} \SubTitle{Software \PFVersion}
\SubTitle{April, 2009} \Author{ 
Reed M. Maxwell\footnote{ \large {\em Department of Geology and Geologic Engineering, Colorado School of Mines, Golden, CO, USA.} \textsf{rmaxwell@mines.edu}}, 
Stefan J. Kollet\footnote{\large {\em Meteorological Institute, Bonn University, Bonn, Germany.} \textsf{stefan.kollet@uni-bonn.de }}, 
Steven G. Smith\footnote{\large {\em Center for Applied Scientific Computing, Lawrence Livermore National Laboratory, Livermore, CA. USA.} \textsf{sgsmith@llnl.gov} }, 
Carol S. Woodward\footnote{\large {\em Center for Applied Scientific Computing, Lawrence Livermore National Laboratory, Livermore, CA, USA.} \textsf{cswoodward@llnl.gov}}, 
Robert D. Falgout\footnote{\large {\em Center for Applied Scientific Computing, Lawrence Livermore National Laboratory, Livermore, CA, USA.}},  
Chuck Baldwin, 
William J. Bosl \footnote{\large {\em Children's Hospital Informatics Program, Harvard Medical School, Boston, MA, USA.}}, Richard Hornung\footnote{\large {\em Center for Applied Scientific Computing, Lawrence Livermore National Laboratory, Livermore, CA, USA.}}, 
Steven Ashby\footnote{\large {\em Pacific Northwest National Laboratory, Richland, WA, USA.}}
}


\end{TitlePage}

%=============== Copyright Page

\begin{CopyrightPage}
\noindent ParFlow is released under the GNU LPGL License \newline
\noindent Version 1.3, 3 November 2008 \newline
\noindent Copyright \copyright{} 2000, 2001, 2002, 2007, 2008  Free Software Foundation, Inc.\newline
\code{http://fsf.org/}\\
This manual is licensed under the GNU Free Documentation License.\newline\vspace{1em}
Copyright \copyright{} 2009  Reed M. Maxwell, Stefan J. Kollet, Steven G. Smith, Carol S. Woodward. 
Permission is granted to copy, distribute and/or modify this document
under the terms of the GNU Free Documentation License, Version 1.3
or any later version published by the Free Software Foundation; with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of the license is included in the section entitled "GNU Free Documentation License".\newline\vspace{1em}
Permission is granted to make and distribute verbatim copies of this
manual provided the copyright notice and this permission notice are
preserved on all copies.\newline\vspace{0.1em}
This computer software and documentation was prepared as an account of
work sponsored by an agency of the United  States Government.  Neither
the United States Government nor the  University of California nor any
of their employees, makes any warranty, express or implied, or assumes
any legal liability  or responsibility for the accuracy, completeness,
or   usefulness of any   information,  apparatus, product, or  process
disclosed,   or represents that  its  use would not infringe privately
owned rights.  Reference  herein to any specific  commercial products,
process,   or service by   trade  name,  trademark,  manufacturer,  or
otherwise, does not  necessarily constitute or imply its  endorsement,
recommendation, or  favoring  by the  United  States Government or the
University of California.  The views and opinions of authors expressed
herein do not necessarily state or reflect those  of the United States
Government or the University of California, and shall  not be used for
advertising or product endorsement purposes.
\end{CopyrightPage}

%=============== Table of Contents

\pagenumbering{roman}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

%=============== Include Chapters

\input{intro}
\input{start}
\input{pfsystem}
\input{models}
\input{files}
\input{fdl}

%=============== Include Other

 %\input{glossary}

%=============== Print the References here

\bibliographystyle{plain}
\bibliography{parflow}

%=============== Print the Index here

\printindex

\end{document}
